---
title: "Quarto Basics"
author: "Aaron M. Geller"
date: "12/12/2023"
number-sections: true
format:
    html:
        code-fold: false
        code-tools: true
theme: sketchy
toc: true
---

## Start in R
```{r}
#| label: r-code-1

library(reticulate)

# define a simple variable as a test
x <- 123

```

## Use variables from R in Python
```{python}
#| label: python-code-1

# multiply the x value from r by 2
y = r.x*2.

```


## Use variables from Python in R
```{r}
#| label: r-code-2

# add to the y variable from python 
z <- py$y + 7
```

## Markdown using variables from both R and Python.

I started in **R** with a value of $x = `r x`$.  Then I used **Python** to define $y = 2 \times x = `r py$y`$.  Then I moved back to **R** to define $z = y + 7 = `r z`$.


## Create a plot in Python using the iris dataset from R.  

I will use  [`seaborn`](https://seaborn.pydata.org/) to generate the corner plot in @fig-corner. (I haven't found a package in R that can produce a nicer looking corner plot in fewer lines of code.) 

### Load the data in R
```{r}
#| label: r-code-3

library(datasets)

# load the iris dataset and print the first 10 rows
data(iris)
head(iris)

```


### Create the plot in Python 


```{python}
#| label: fig-corner
#| fig-cap: "Corner plot created with seaborn showing iris flower data."

# make a plot of the iris data
import seaborn as sns

p = sns.PairGrid(r.iris, diag_sharey = False, hue = 'Species', corner = True).map_lower(sns.scatterplot).map_diag(sns.kdeplot).add_legend(bbox_to_anchor = (0.6, 0.6))

# for RStudio you should include the line below (or you may not see the plot)
# p

```

## New Python section
```{python}
#| label: python-code-2

# add all the variables together
a = r.x + y + r.z

```

Let's check this math : $a = x + y + z = `r x` + `r py$y` + `r z` = `r py$a`$.